 

### 2021-06-11 vertical-thermal-review.Rmd
Update vertical-thermal-review.Rmd

Leaf respiration is lower during the day than the night, due to light-induced inhibition that is highest in high-irradiance and moderate-to-high-temperature and lowest at high-irradiance and low-temperature [@atkinLeafRespirationSnow2000; @keenanWidespreadInhibitionDaytime2019; @zaragoza-castellsDoesGrowthIrradiance2007]. Light-induced inhibition of leaf respiration influences $Q_{10}$ where $Q_{10}$ is greatest in the dark, and decreases with increase in irradiance [@atkinLeafRespirationSnow2000].

However, when it comes to sun and shade leaves, a few studies have found $Q_{10}$ to be greater in upper-canopy leaves than in shaded lower-canopy leaves across species [@turnbullScalingFoliarRespiration2003; @harleyEffectsLightTemperature1996] and other studies have observed no difference in $Q_{10}$ across species vertical profiles and seedling sun vs. shade leaves [@weerasingheCanopyPositionAffects2014; @bolstad_foliar_1999; @zaragoza-castellsDoesGrowthIrradiance2007; @zaragoza-castellsClimatedependentVariationsLeaf2008], and at Coweeta, temperate trees showed a slight tendency to increase down the canopy relative to upper-canopy leaves as $T_{air}$ increased [@bolstad_foliar_1999](*see their Fig. 1*). (*Martijn has some unpublished data on this that we might include.unpublished data showing higher Q10 in shade than in sun leaves for 10 species or so. May make sense to include here: https://github.com/EcoClimLab/vertical-thermal-review/issues/8.*). @bolstad_foliar_1999 also found evidence of acclimation, but there remained a declining trend of respiration with elevation.
Furthermore, response of respiration to $T_{leaf}$ [partially represented  by $E_{0}$] showed no variation in canopy positions of Quercus rubra, whereas on considering site differences, canopy-position effect in respiration was much greater in mesic sites, and seasonal effect was greater in drier sites [@xuSeasonalVariationTemperature2006].

>>>>

A few studies have found $Q_{10}$ to be greater in upper-canopy leaves than in shaded lower-canopy leaves, with significant differences among species [@turnbullScalingFoliarRespiration2003; @harleyEffectsLightTemperature1996], whereas Griffin et al. 2002 found that in Populas deltoides, lower canopy leaf respiration was more sensitive to changes in $T_{air}$ than upper canopy. Other studies have observed no difference in $Q_{10}$ across species vertical profiles and seedling sun vs. shade leaves [@weerasingheCanopyPositionAffects2014; @bolstad_foliar_1999; @zaragoza-castellsDoesGrowthIrradiance2007; @zaragoza-castellsClimatedependentVariationsLeaf2008], although at Coweeta, temperate trees showed a slight tendency to increase down the canopy relative to upper-canopy leaves as $T_{air}$ increased [@bolstad_foliar_1999](*see their Fig. 1*). (*Martijn has some unpublished data on this that we might include.unpublished data showing higher Q10 in shade than in sun leaves for 10 species or so. May make sense to include here: https://github.com/EcoClimLab/vertical-thermal-review/issues/8.*). @bolstad_foliar_1999 also found evidence of acclimation, but there remained a declining trend of respiration with elevation.
Furthermore, response of respiration to $T_{leaf}$ [partially represented  by $E_{0}$] showed no variation in canopy positions of Quercus rubra, whereas on considering site differences, canopy-position effect in respiration was much greater in mesic sites, and seasonal effect was greater in drier sites [@xuSeasonalVariationTemperature2006].
